Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-26 abstract/abstract.tex
Updates from Overleaf

\bigskip
Keywords: Music; Music Theory; Music Information Retrieval; Music Structure Analysis; Machine Learning; Neural Networks; Deep Learning

>>>>

\bigskip
Keywords: Music; Music Theory; Music Information Retrieval; Music Structure Analysis; Machine Learning; Neural Networks; Deep Learning; Self-Supervised Learning; Music Similarity; Epidemic Sound

### 2023-03-26 introduction/Subsections/Introduction.tex
Updates from Overleaf

Humanity has been studying music to understand its underlying structure for millennia. From the ancient Greeks, who believed that music had a mathematical and cosmic harmony, to the early works of theorists such as Rameau, Fux, and Schenker, and more recent approaches such as generative theory and corpus studies, music has been a subject of fascination and inquiry. This rich music theory and analysis history provide a valuable foundation for guiding and evaluating MIR algorithms and models.

@@

Music Information Retrieval (MIR) is a research area that has emerged in the last decades as an interdisciplinary field combining engineering, musicology, and neuroscience. MIR aims to develop algorithms and techniques to automatically analyze, classify, and retrieve musical information from audio signals. In my opinion, MIR research has been traditionally focused on technical aspects, such as digital signal processing (DSP), machine learning, and pattern recognition, neglecting other relevant domains, such as perception, cognition, music theory, aural skills, and musicology.

The reason behind this is non-other but the high implementation complexity for such heuristics and abstract fields of study.

>>>>

Humanity has been studying music to understand its underlying structure for millennia. From the ancient Greeks to the early days, music has been a subject of fascination and inquiry. This rich music theory and analysis history provide a valuable foundation for guiding and evaluating Music Information Retrieval (MIR) algorithms and models.

@@

MIR is a research area that has emerged in the last decades as an interdisciplinary field combining engineering, musicology, and neuroscience. MIR aims to develop algorithms and techniques to automatically analyze, classify, and retrieve musical information from mainly audio signals. I believe that traditionally, MIR research has primarily concentrated on technical aspects, such as Digital Signal Processing (DSP), or pattern recognition. As a result, other relevant domains, including perception, cognition, music theory, aural skills, and musicology, along with its subfields, such as the symbolic domain, orchestration, and automatic transcription, have often been overlooked. However, it is worth noting that there has been a significant shift in this tendency recently.

The reason behind this is non-other but the implementation complexity for such heuristics and abstract fields of study.

With the advent of Machine Learning (ML), MIR has expanded to encompass musically motivated neural network architectures\cite{musicmotivCNN}. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

### 2023-03-26 introduction/Subsections/Motivation.tex
Updates from Overleaf

>>>>

\section{About aural skills and high-level perceptual}

%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees, and the tonal function.}
\label{fig:timeseries}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

I have consistently been inspired by researchers who strive to uncover musical ground truth within the existing tonal paradigm, such as Heinrich Schenker\cite{}, or by challenging it, as exemplified by George Russell\cite{LydianRussell}. Both approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth and development.

Schenkerian Analysis:
Schenkerian Analysis, developed by Austrian music theorist Heinrich Schenker, analyzes tonal music focusing on its underlying structure. It seeks to reveal the hierarchical relationships between pitches, based on the premise that all tonal music has a fundamental structure known as the "Ursatz" or "basic shape." The Ursatz consists of a stepwise descending line (Urlinie) supported by a bass arpeggiation (Bassbrechung). Schenkerian Analysis reduces a musical composition to its essential elements, enabling the analyst to observe how the work's various layers contribute to the overall structure. It is worth mentioning that such analysis is done in the symbolic domain.

George Russell's Lydian Chromatic Concept:
George Russell's Lydian Chromatic Concept of Tonal Organization is an innovative approach to understanding and organizing musical relationships. Introduced in the 1950s, the theory asserts that the Lydian mode, rather than the traditional major scale, is the accurate parent scale in Western music. The Lydian scale is characterized by a raised fourth degree, giving it a brighter and more consonant sound. Russell's concept revolves around "chord modes," derived from a chord's parent Lydian scale. This system allows for greater harmonic flexibility and a broader range of tonal colors, influencing the development of jazz and contemporary music.

### 2023-03-26 introduction/Subsections/Objectives.tex
Updates from Overleaf

With the advent of Machine Learning (ML), MIR has expanded to encompass musically motivated neural network architectures. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

By studying music structure analysis using musically motivated neural networks, we will gain an in-depth understanding of the latest advancements in the field and how these techniques are being used to revolutionize our understanding of music. Additionally, we will be able to contribute to developing new and innovative methods for musical analysis.

This work aims to generalize the fundamental tonal-working structure of western-tradition music by learning and embedding space used to facilitate its application in downstream MIR tasks. Understanding how music is organized and composed is critical, and music theory plays a vital role.

>>>>

This work aims to uncover a generalization of the fundamental tonal-working structure of western-tradition music by learning and embedding space used to facilitate its application in downstream MIR tasks.

Understanding how music is organized and composed is critical, and music theory plays a vital role.
